---
title: "MBTA project EDA"
author: "runze"
date: "2022-12-18"
output: pdf_document
---

```{r setup, include=FALSE}
library(readr)
library(mapsapi)
library(leaflet)
library(vroom)
library(lubridate)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(tibble)
library(shiny)
```

```{r}
setwd("E:/MBTAproj")
#load stops info (lat-lon)
stops<-vroom("stops.csv")
stops<-stops[complete.cases(stops$stop_code),]
LRQ4_21<-vroom("LRTravelTimesQ4_21.csv")
LRQ1_22<-vroom("2022-Q1_LRTravelTimes.csv")
LRQ2_22<-vroom("2022-Q2_LRTravelTimes.csv")
LRQ3_22<-vroom("2022-Q3_LRTravelTimes.csv")

T_2122<-rbind(LRQ4_21,LRQ1_22,LRQ2_22,LRQ3_22)

rm(LRQ4_21)
rm(LRQ1_22)
rm(LRQ2_22)
rm(LRQ3_22)


T_2122$service_date<-as.Date(T_2122$service_date)
T_2122$month<-month(T_2122$service_date)
T_2122$week<-week(T_2122$service_date)
T_2122$weekdays<-weekdays(T_2122$service_date,abbreviate = T)

#Randomly choose a week in these 12 months
sample_weeks<-c()
set.seed(726)
for (i in 1:12) {
  weeks_in_month<-levels(as.factor(T_2122$week[T_2122$month==i]))
  sample_weeks<-c(sample_weeks,sample(weeks_in_month[2:(length(weeks_in_month)-1)],1)) # To make sure pick out an entire week in each month, I exclude the fist week and last week in each month.
}


smp_T<-T_2122[T_2122$week %in% as.numeric(sample_weeks),]

rm(T_2122)

smp_T$trip<-paste0(paste0("From ",smp_T$from_stop_id),paste0(" to ",smp_T$to_stop_id))# Add a variable "trip" 

smp_T$trip<-as.factor(smp_T$trip)

```



# The light rail data EDA

 Check the overall distribution of travel times

```{r}
summary(smp_T$travel_time_sec)
hist(smp_T$travel_time_sec,breaks = 120,main = "Overall distribution of travel time (in seconds)",xlab = "Travel time")
```

 Check the distribution of travel times for each weekdays

```{r}
ggboxplot(smp_T,x = "weekdays",y="travel_time_sec")
```

 We can check many kinds of trips are there

```{r}
trip<-levels(smp_T$trip)

length(trip)

```
 There are total 2067 kinds of trips for light rails in boston

 We can check which trip has the highest frequency

```{r}
table(smp_T$trip)[table(smp_T$trip)==max(table(smp_T$trip))]
```
 The most busy trip is from station 70157 to station 70155, which has a frequency of 39928

 Then check how this trip's travel times are distributed

```{r}
barplot(smp_T$travel_time_sec[smp_T$trip=="From 70157 to 70155"],ylab = "Travel time")
```
 We can see that for most cases, this trip finishes within 200 seconds, but for some extream situation, the travel time can up to 600 seconds.








